### 2015-05-08 evaluation.tex
updates to evaluation
added draft plots in evaluation section
update in evaluation

\begin{table*}[t]
\small
\centering
\setlength{\tabcolsep}{0.5em}
\caption{Comparison with different yield instrumentation granularities and context-switch bounds.}
\label{tab:results}
\begin{tabular}{l r rrrr rrrr rrrr}
\centering
\input{experiments/tables/scalability.tex}
\end{tabular}
\end{table*}

@@

\noindent
\textbf{Benchmarks}\xspace\xspace We evaluate our methodology against \sizeOfBenchmarks drivers taken from the 4.0 Linux kernel. We chose drivers from many different domains: block; char; ethernet; near field communication (nfc); universal serial bus (usb); and watchdog. Further, we focused on choosing non-trivial drivers (e.g. large number of entry points, deep nesting of entry point helper functions, complex use of locking, use of function pointers, etc). We manually modeled the environment for all these drivers, a process that required more than a month of work.

\noindent
\textbf{Race-Checking}\xspace\xspace Table~\ref{tab:stats} presents statistics for all our benchmarks: lines of code (LoC); number of entry point pairs (\#Pairs); number of SMACK memory regions (\#MRs); number of racy pairs that \whoop identified (\#Racy Pairs); number of racy memory regions that \whoop reported (\#Racy MRs); and number of races that \corral discovered (\#Races Found) using a context-switch bound of 2.

We can see that, in general, \whoop reports more races than \corral. This is expected because \whoop over-approximates the original driver for soundness: if a race exists, the tool will report it (assuming tool and model correctness), but \whoop can, and does, report false races. On the contrary, \corral is precise, but can miss races as it explores the driver under pre-defined bounds. Further, a common problem for \corral is loop coverage. To tackle this we enable the build-in loop over-approximation described in~\cite{lal2014powering}. This can potentially lead \corral to report false positives, but we have not seen this in practice.

Most of the data races that \whoop and \corral found can be classified in two cases. The first case had to do with incrementing a global counter without using a lock. The reason behind this might be performance, but even benign races can lead to undefined behavior according to the C standard. The second case had to do with an entry point modifying a field of an object (either global or passed as parameter) without using a lock. This can easily lead to a data race if another concurrently running entry point accesses the same field of the same object. As an example of the second case, we discovered the following data race in the generic\_nvram char driver: during the \texttt{llseek} entry point, the driver is accessing the file offset \texttt{file->f\_pos} (\texttt{file} is passed as a parameter to the entry point) without first acquiring a lock. This can lead to a race if the driver invokes \texttt{llseek} from another thread passing the same \texttt{file} object as a parameter. When we investigated another char driver that uses the same API, we saw that it was protecting the offset access in the same entry point with a mutex. This made us suspicious that the race we found in generic\_nvram is real. We filed a bug report, but did not manage to get a response yet.

It is worthwhile to mention that we did not discover any additional data races with larger context switch bounds. This might mean that most races in these drivers require only a small amount of context-switches to manifest, or that \corral run out of resources before discovering the deeper bug.

>>>>

\begin{table*}[t]
\small
\centering
\setlength{\tabcolsep}{0.45em}
\caption{Comparison with different yield instrumentation granularities and context-switch bounds.}
\label{tab:results}
\begin{tabular}{l r rrrr rrrr rrrr}
\centering
\input{experiments/tables/scalability.tex}
\end{tabular}
\end{table*}

\begin{figure}
\centering
\includegraphics[width=.99\linewidth]{experiments/figures/yieldmr_vs_yieldall.pdf}
\caption{draft}
\label{fig:plot1}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=.99\linewidth]{experiments/figures/correlation.pdf}
\caption{draft}
\label{fig:plot2}
\end{figure}

@@

\noindent
\textbf{Benchmarks}\xspace\xspace We evaluate our methodology against \sizeOfBenchmarks drivers taken from the 4.0 Linux kernel. We chose non-trivial drivers from many different domains: block; char; ethernet; near field communication (nfc); universal serial bus (usb); and watchdog. We had to manually model the environment for these drivers, a process that required two months of work.

\noindent
\textbf{Race-Checking}\xspace\xspace Table~\ref{tab:stats} presents statistics for all our benchmarks: lines of code (LoC); number of entry point pairs (\#Pairs); number of SMACK memory regions (\#MRs); number of racy pairs that \whoop identified (\#Racy Pairs); number of racy memory regions that \whoop reported (\#Racy MRs); and number of races that \corral discovered (\#Races Found) using a context-switch bound of 2. We did not discover new races with larger context switch bounds. This might mean that these drivers only have races that manifest with at least one or two context-switches, or that \corral run hit its bounds before discovering a deeper race.

We can see that \whoop reports more races than \corral. This is expected, as our approach uses over-approximation (to find all races), and thus can introduce false bugs. On the other hand, \corral is precise, but can miss races as it explores programs under pre-defined bounds. Another issue with \corral is loop coverage due to unsound loop unrolling. To tackle this, we enable the built-in loop over-approximation described in~\cite{lal2014powering}. This can potentially lead \corral to report false positives, but we have not seen this in practice.

Most of the races that \whoop and \corral discovered can be classified in two cases. The first case is accessing a global counter (or flag) from concurrent entry points, without a lock. This might be for performance, and indeed a lot of the races we found might be benign. Even benign races, though, can lead to undefined behavior according to the C standard, which can cause many issues (e.g. with compilers). The second case is an entry point accessing a field of an object (either global or passed as a parameter) without a lock. This can lead to a race if another entry point simultaneously accesses the same field of the same object.

As an example of the second case, we discovered the following race in the generic\_nvram char driver: during the \texttt{llseek} entry point, the driver is accessing the file offset \texttt{file->f\_pos} without first acquiring a lock (\texttt{file} is passed as a parameter to \texttt{llseek}). This can lead to a race if the driver invokes \texttt{llseek} from another thread passing the same \texttt{file} object as a parameter. When we investigated another char driver that uses the same APIs, we saw that it was protecting the offset access in its \texttt{llseek} entry point with a mutex. This made us suspicious that the race we found in generic\_nvram could be real. We filed a bug report, but did not manage to get a response yet.
